---
title: "Preliminary Corvid results for King County/Seattle public schools"
author: "Dennis Chao (dennisc@idmod.org, Institute for Disease Modeling)"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[RO]{\today}
- \fancyfoot[CO]{IDM}
- \fancyfoot[R]{\thepage}
output:
  pdf_document: 
    fig_caption: yes
    fig_width: 6.5
    fig_height: 5.5
---

```{r initialize, echo=FALSE}
library(knitr)
library(RColorBrewer)
palr0 <- c("black", "red", "red")
names(palr0) <- c("2.6","2.0", "2")

palage <- c("lightblue", "blue", "tan", "brown", "red")
nameage <- c("<5y", "5-18y", "19-29", "30-64", "65+")
durations <- c(14,28,56,140)
```

# Introduction

We ran school closure Scenarios in response to a request by King County using Corvid, a newly developed  agent-based model of COVID-19 transmission.
The Corvid model was based on FluTE (Chao et al 2010), which was based on
the EpiCast model described in Germann et al 2006.
We describe the Corvid model near the end of this document.
In the Results, we run the model on a synthetic population based on 2000 Census data for metropolitan Seattle.
We consider this population represents a moderately large American city, though it is not an accurate representation of present-day Seattle.

In general, school closures alone delayed but did not reduce the epidemic peak.
The peak could be reduced when school closures spanned the peak of the epidemic, which requires either very long school closures or rather late school closures that start while the epidemic has already ramped up.

Other social distancing policies, such as liberal leave and work from home policies, may have a more durable effect since we assume workplaces that comply with these policies can do so for the duration of the epidemic (workplaces do not re-open during the epidemic in our simulations).
Combining both school closures and workplace-enacted social distancing policies can delay and reduce the epidemic peak.
However, Corvid is a new model under active development, so results must be interpreted with caution.
These results suggest that sustainable interventions be enacted widely to reduce transmission, while school closures can be used to "buy time".

We believe that a combination of social distancing policies should and will be enacted to slow the spread of COVID-19.
In the Corvid model, we implemented "liberal leave" (workers have an elevated probability of staying home when sick) and "work from home" (workers stay home instead of going to work but still have some interaction with people in their home community).
We ran simulations in which 50% of the adult population complied with these measures, though the real number might be lower since many adults will not have these options available.
We believe that these other social distancing policies will be implemented on an ad-hoc basis, and school closure may force some parents to stay home anyways (de facto "work from home").
We believe that school closures generally delay the peak, but do not usually reduce the peak or total cases because they are temporary.
More sustainable measures, like workplace liberal leave and work from home policies, can flatten the epidemic curve and reduce the total number of cases.
Combining school closures with these other social policies may be synergistic, but we only tested a limited number of scenarios (one).
With concurrent effective policies enacted, the epidemic can be "suppressed."

\clearpage

# Results

```{r load-seattle, echo=FALSE}
tractlocs <- read.csv("../corviddata/seattle-tracts.dat", header=FALSE, col.names=c("fipsstate", "fipscounty", "fipstract", "population", "latitude", "longitude"))
tracts.seattle <- read.csv("results/Tracts-seattle20-60-school-none-0-llwfh-0.txt", header=TRUE)
tracts.seattle$pop <- rowSums(tracts.seattle[5:9])
popsize.seattle <- sum(tracts.seattle$pop)

# load the "no intervention" outputs
symptomatic.seattle <- list()
csymptomatic.seattle <- list()
newlysymptomatic.seattle <- list()
for (r0 in c(2.0,2.6)) {
#  temp <- readLines(paste("results/Summary-seattle",r0*10,"-60-school-none-0-llwfh-0.txt",sep=""))
  temp <- readLines(paste("results/Summary-seattle",r0*10,".txt",sep=""))
  symptomatic.seattle[[as.character(r0)]] <- as.numeric(strsplit(gsub("^.*: ","",temp[grep("Number symptomatic", temp)]),",")[[1]])
  csymptomatic.seattle[[as.character(r0)]] <- as.numeric(strsplit(gsub("^.*: ","",temp[grep("Cumulative symptomatic", temp)]),",")[[1]])
  newlysymptomatic.seattle[[as.character(r0)]] <- c(symptomatic.seattle[[as.character(r0)]][1], diff(csymptomatic.seattle[[as.character(r0)]]))
}
```

We show sample plots with no interventions to show the dynamics of the epidemic for R0=2.0 and 2.6.
Finally, we test combining other social distancing measures with school closure.
Population of `r popsize.seattle`.
Seeded 5 people.


## No intervention

We initialize the model by infecting five people on day 0, then run the model for 180 simulated days.
We ran the model with R_0=2.0 and 2.6.

```{r cases-seattle, fig.width=7, fig.height=3.25, fig.cap="Symptomatic people over time in Seattle with no interventions. On the right, cumulative symptomatic people over time. The axis on the right is the percent of the total population.", echo=FALSE}
par(mfrow=c(1,2),
    mar=c(3.5,3.5,1,2), #bottom, left, top, and right.
    mgp=c(2.0, 0.6, 0))
ymax <- 7000
plot(x=NA, y=NA, xlim=c(0,180), ylim=c(0,ymax), xlab="days since introduction", ylab="newly symptomatic per day", type="l", col=palr0["3.0"], lwd=2, cex.axis=0.8)
for (r0 in c(2.0,2.6)) {
    days <- 1:length(newlysymptomatic.seattle[[as.character(r0)]])
  lines(x=days, y=newlysymptomatic.seattle[[as.character(r0)]], lwd=2, col=palr0[as.character(r0)])
}
legend("topleft", legend=c("R0=2.6", "R0=2.0"), fill=palr0[c("2.6","2.0")], inset=0.01, cex=0.8)

ymax <- popsize.seattle*0.45
plot(x=NA, y=NA, xlim=c(0,180), ylim=c(0,ymax), xlab="days since introduction", ylab="cumulative symptomatic", type="l", col=palr0[as.character(r0)], lwd=2, cex.axis=0.8)
axis(4, at=sum(tracts.seattle$pop) * seq(0,100,10)/100, lab=paste(seq(0,100,10),"%",sep=""), las=2, col="gray", col.axis="darkgray")
for (r0 in c(2.0,2.6)) {
  lines(x=1:length(csymptomatic.seattle[[as.character(r0)]]), y=csymptomatic.seattle[[as.character(r0)]], lwd=2, col=palr0[as.character(r0)])
}
legend("topleft", legend=c("R0=2.6", "R0=2.0"), fill=palr0[c("2.6","2.0")], inset=0.01, cex=0.8)
```

\clearpage
## Effect of school closures and other social distancing policies

We believe that a combination of social distancing policies should and will be enacted to slow the spread of COVID-19.

When schools close in Corvid, children no longer transmit in schools, but they do interact with anyone present in their households during the day as well as a low level of interaction with everyone in their local community.
In this section, we assume that school closures are implemented without other social distancing measures.
In the Corvid model, we implemented two workplace policies:  "liberal leave" (workers have an elevated probability of staying home when sick) and "work from home" (workers stay home instead of going to work but still have some interaction with people in their home community).
For now, we are setting the compliance for both these workplace measures to be the same.

We ran simulations in which 25% of the adult population complied with these measures, though the real number might be lower since many adults will not have these options available or higher because of anxiety.
We believe that these other social distancing policies will be implemented on an ad-hoc basis, and school closure may force some parents to stay home anyways (de facto "work from home").
We believe that school closures generally delay the peak, but do not usually reduce the peak or total cases because they are temporary.
More sustainable measures, like workplace liberal leave and work from home policies, can flatten the epidemic curve and reduce the total number of cases.
Combining school closures with these other social policies may be synergistic, but we only tested a limited number of scenarios (one).

Below we plot the impact of starting combinations of school closure and
the workplace policies (liberal leave and work from home) on days 60, 74, and 100.
We plot the number of sick people over time below for our synthetic Seattle population.

The greatest impact of school closures is on reducing the peak of the epidemic, which can alleviate strain on hospitals.
But school closures must be in effect right before cases peaking for maximum reduction of the peak of both cases and hospitalizations.
Below are tables of the numbers of deaths, hospital demand (total hospital occupancy days), and peak hospital use (in units of "per 100,000 general population")
for all combinations of R0, duration of school closure (0 for no closure), threshold for starting the school closures.
These are followed by plots showing the relationship between total hospital demand vs hospital peak demand.

There are some non-intuitive results concerning the impact of school closures.
If the threshold at which you start school closures is very low (0.0001), then schools are shut down early and re-open before the epidemic dies out, which means there is almost no reduction in hospitalizations or deaths.
Basically, the interventions end too early in the course of the outbreak when you start "too early".
Almost any long (more than 28 days) early school closure seems to delay the peak.

```{r cases-seattle-socialdistancing, fig.width=7, fig.height=8.25, fig.cap="Impact of social distancing, R0=2.6. The dashed vertical line is the start of social distancing policies, including school closures, liberal leave, and work from home. We test school closures alone, work from home/liberal leave (25% compliance) without school closures, and a combination of all three policies. Triangles near the x-axis mark when the schools re-open. The workplace policies (liberal leave and working from home) continue indefinitely. On the left, daily new symptomatic cases per 100,000 population. On the right, cumulative % of the population symptomatic.", echo=FALSE}
r0 <- 2.6
duration <- 42 # school closures
denominator <- 100000 # scale results
scenarios <- c("school closure only", "wfh/ll only", "school closure + wfh/ll")
palscenario <- c("black","red","orange","blue")
names(palscenario) <- c("no intervention",scenarios)
par(mfrow=c(3,2),
    mar=c(3.5,3.5,2,2.2), #bottom, left, top, and right.
    mgp=c(2.0, 0.6, 0))
days <- 1:length(newlysymptomatic.seattle[[as.character(r0)]])
    compliance <- 25
for (startday in c(60,74,100)) {
    ymax <- 1200
    plot(x=days, y=newlysymptomatic.seattle[[as.character(r0)]]*denominator/popsize.seattle, xlab="days since introduction", ylab=paste("symptomatic /", as.integer(denominator)), main=paste("R0=",r0,", start day=",startday, ", compliance=",compliance,"%",sep=""), type="l", col="black", lwd=2, cex.main=1.3, cex.axis=1.1, cex.lab=1.2, xlim=c(0,180), ylim=c(0,ymax))
    polygon(x=c(days,max(days), 0), y=c(newlysymptomatic.seattle[[as.character(r0)]],0,0)*denominator/popsize.seattle, col="gray92", lwd=0.5)
    lines(x=rep(startday,2), y=c(-100,ymax*1.1), lty="dashed")    
    symp <- list()
    csymp <- list()
    nsymp <- list()
    for (scenario in scenarios) {
        schoolpolicy <- ifelse(grepl("school",scenario),"all","none") # close schools or not?
        wfhcompliance <- ifelse(grepl("wfh",scenario),compliance,0)
        temp <- readLines(paste("results/Summary-seattle",r0*10,"-",startday,"-school-",schoolpolicy,"-",ifelse(schoolpolicy=="none",0,duration),"-llwfh-",wfhcompliance,".txt",sep=""))
        symp[[scenario]] <- as.numeric(strsplit(gsub("^.*: ","",temp[grep("Number symptomatic", temp)]),",")[[1]])
        csymp[[scenario]] <- as.numeric(strsplit(gsub("^.*: ","",temp[grep("Cumulative symptomatic", temp)]),",")[[1]])
        nsymp[[scenario]] <- c(symp[[scenario]][1], diff(csymp[[scenario]]))
        lines(x=1:length(nsymp[[scenario]]), y=nsymp[[scenario]]*denominator/popsize.seattle, col=palscenario[scenario], lwd=1.5)
        points(x=duration+startday, y=0, pch=25, bg="red", cex=1.5) # school re-opening date
    }
    legend("topleft", legend=c("nothing", scenarios), fill=c(palscenario), cex=0.75, inset=0.01, bg="white")

    ymax <- 45
    plot(x=1:length(csymptomatic.seattle[[as.character(r0)]]), y=csymptomatic.seattle[[as.character(r0)]]*100/popsize.seattle, xlab="days since introduction", ylab="cumulative symptomatic, %", main=paste("R0=",r0,", start day=",startday, sep=""), type="l", col="black", lwd=2, cex.main=1.3, cex.axis=1.1, cex.lab=1.2, xlim=c(0,180), ylim=c(0,ymax))
    lines(x=rep(startday,2), y=c(-100,ymax*1.1), lty="dashed")
    for (scenario in scenarios) {
        lines(x=1:length(csymp[[scenario]]), y=csymp[[scenario]]*100/popsize.seattle, col=palscenario[scenario], lwd=1.5)
        points(x=duration+startday, y=0, pch=25, bg="red", cex=1.5) # school re-opening date
    }
    legend("topleft", legend=c("nothing", scenarios), fill=c(palscenario), cex=0.75, inset=0.01, bg="white")
}
```

\clearpage
## Reducing the epidemic peak with school closures

```{r cases-seattle-peak, fig.width=7, fig.height=6.5, fig.cap="Timing school closures to reduce the epidemic peak. School closures with different start dates and duration were simulated, shown in different colors. The days when schools are closed are shown as color-coded horizontal lines near the x-axis. The top left plot shows number of newly symptomatic people per day per 100,000 population, and the top right shows cumulative % of the population symptomatic. The bottom plot is R_effective over time. Dashed gray horizontal lines are at 2.6 (R0) and 1.0. One stochastic run per scenario.", echo=FALSE}
r0 <- 2.6
denominator <- 100000 # scale results to standard pop size
par(mfrow=c(2,2),
    mar=c(3.3,3.3,2.5,1), #bottom, left, top, and right.
    mgp=c(2.35, 0.6, 0))
palscenario <- c("black","firebrick","seagreen","orangered","royalblue")
filenames <- c("Summary-seattle26-1-school-all-365-llwfh-0-runlength240.txt",
               "Summary-seattle26-60-school-all-365-llwfh-0-runlength240.txt",
	       "Summary-seattle26-60-school-all-42-llwfh-0-runlength240.txt",
	       "Summary-seattle26-90-school-all-42-llwfh-0-runlength240.txt")
days <- 1:length(newlysymptomatic.seattle[[as.character(r0)]])
for (plottype in c("symptomatic", "cumulative")) {    
    ymax <- ifelse(plottype=="cumulative", 42, 1200)
    y <- newlysymptomatic.seattle[[as.character(r0)]]*denominator/popsize.seattle
    if (plottype=="cumulative") {
        y <- 100*csymptomatic.seattle[[as.character(r0)]]/popsize.seattle
    }
    plot(x=days, y=y, xlab="days since introduction", ylab=ifelse(plottype=="cumulative", "cumulative symptomatic, %", paste("symptomatic /", as.integer(denominator))), main=paste(ifelse(plottype=="cumulative", "Cumulative symptomatic", "Newly symptomatic per day"), ", R0=",r0,sep=""), type="l", col="black", lwd=2, cex.main=1.1, cex.lab=1.1, ylim=c(ifelse(plottype=="cumulative",-4,-100),ymax), axes=FALSE)
    axis(1, at=seq(0,365,30), cex.axis=0.8)
    if (plottype=="cumulative") {
        axis(2, at=seq(0,100,10), lab=paste(seq(0,100,10), "%", sep=""), cex.axis=0.9, las=2)
    } else {
        axis(2, cex.axis=0.9)
    }    
    polygon(x=c(days,max(days), 0), y=c(y,0,0), col="gray92", lwd=0.2)
    symp <- list()
    csymp <- list()
    nsymp <- list()
    responseday <- vector()
    closureduration <- vector()
    for (filenum in 1:length(filenames)) {
	scenario <- filenames[filenum]
        temp <- readLines(paste("results/",filenames[filenum],sep=""))
        symp[[filenum]] <- as.numeric(strsplit(gsub("^.*: ","",temp[grep("Number symptomatic", temp)]),",")[[1]])
        csymp[[filenum]] <- as.numeric(strsplit(gsub("^.*: ","",temp[grep("Cumulative symptomatic", temp)]),",")[[1]])
        nsymp[[filenum]] <- c(symp[[filenum]][1], diff(csymp[[filenum]]))
        responseday[filenum] <- as.numeric(strsplit(gsub("^.*: ","",temp[grep("Reactive strategies deployed on day", temp)]),",")[[1]])
        closureduration[filenum] <- as.numeric(strsplit(gsub("^.*: ","",temp[grep("School closure days", temp)]),",")[[1]])
	if (plottype=="cumulative") {
            lines(x=1:length(csymp[[filenum]]), y=100*csymp[[filenum]]/popsize.seattle, col=palscenario[filenum+1], lwd=2)
	} else {
            lines(x=1:length(nsymp[[filenum]]), y=nsymp[[filenum]]*denominator/popsize.seattle, col=palscenario[filenum+1], lwd=2)
	}
        lines(x=responseday[filenum]+c(0,closureduration[filenum]), y=rep(-filenum*ifelse(plottype=="cumulative",1,30),2), col=palscenario[filenum+1], lwd=4) # school closure dates
    }
    if (plottype=="cumulative") {
        legend("topleft", legend=c("no closures", paste("closed ", closureduration, "days starting day", responseday)), fill=c(palscenario), cex=0.56, inset=0.01, bg="white")
    }
#    legend("topleft", legend=c("no closures", paste("closed ", closureduration, "days starting day", responseday)), fill=c(palscenario), cex=0.3749, inset=0.01, bg="white")
}

# R effective over time
r0 <- 2.6
filenames <- c("Individuals-seattle26.txt",
               "Individuals-seattle26-1-school-all-365-llwfh-0-runlength240.txt",
               "Individuals-seattle26-60-school-all-365-llwfh-0-runlength240.txt",
	       "Individuals-seattle26-60-school-all-42-llwfh-0-runlength240.txt",
	       "Individuals-seattle26-90-school-all-42-llwfh-0-runlength240.txt")
palscenario <- c("black","firebrick","seagreen","orangered","royalblue")
ymax <- 3.6
binsize <- 15
xmax <- 239
    plot(x=NA, y=NA, xlab="days since introduction", ylab="R_effective", main=paste("R_effective over time, R0=",r0,sep=""), type="l", col="black", lwd=2, cex.main=1.1, cex.lab=1.1, xlim=c(0,xmax), ylim=c(-0.5,ymax), axes=FALSE)
    axis(1, at=seq(15,365,30), cex.axis=1, lwd=0.9, tck=-0.02, col="darkgray", col.axis=NA)
    axis(1, at=seq(0,365,30), cex.axis=0.8)
    axis(2, cex.axis=1, las=2)
    abline(a=1,b=0,lty="dashed",col="darkgray")
    abline(a=r0,b=0,lty="dashed",col="darkgray")

    for (filenum in 1:length(filenames)) {
        individuals <- read.csv(paste("results/",filenames[filenum],sep=""), header=TRUE)
	windowstarts <- seq(0,xmax,binsize) # start of window in days
	y <- sapply(windowstarts, function(windowstart) {
	    valid <- individuals$infectedtime/2>=windowstart & individuals$infectedtime/2<windowstart+binsize
	    sum(individuals$sourceid %in% individuals$id[valid])/sum(valid)
	    })
	points(x=windowstarts+binsize/2, y=y, pch=filenum, col=palscenario[filenum], lwd=1)
	lines(x=windowstarts+binsize/2, y=y, col=palscenario[filenum], lwd=3)
	if (filenum>1) {
          lines(x=responseday[filenum-1]+c(0,closureduration[filenum-1]), y=rep(-filenum*0.1,2), col=palscenario[filenum], lwd=4) # school closure dates
	}
    }
    legend("topright", legend=c("no closures", paste("closed ", closureduration, "days starting day", responseday)), col=c(palscenario), pch=1:length(filenames), lwd=1.5, cex=0.75, inset=0.01, bg="white")
```

We believe that you can lower the peak of an epidemic by closing schools for a few weeks around peak, as we noted above.
Here, we model school closures (with no other interventions) with different start times and durations to illustrate this.

In the plot above, we can see that closing schools for 42 days when the epidemic starts ramping up simply pushes the epidemic peak back (compare black and orange curves).
If you never open schools throughout the entire epidemic, from when the first case was infected on day 1 of the epidemic for a whole year, the peak is obviously reduced and pushed back (dark red curve).
If you close schools a little later in the epidemic on day 60 and never re-open them, the peak is not delayed much (green curve), but it is reduced just as much as if school closures had started on day 1 (red curve) and the impact on total cases is about the same (right plot, green vs red curves).
If you close schools for only 42 days a couple of weeks really late, right before the epidemic peaks, you do not delay the peak at all (blue curve), but it is reduced nearly as much as if you had never opened schools at all.
Closing schools late in the epidemic and keeping them closed through the peak may prevent a "bounceback" of transmission when schools reopen.
Though the peak is reduced by this late shorter school closure, it does not reduce the total number of cases (right plot, blue curve vs black curve).

For these school closure scenarios, we plotted R_effective over time.
This can help us understand why the epidemic peaks when it does under different interventions.
For a set of individuals, we computed R_effective taking the number of people infected by that set and dividing by the number in that set. We plotted R_effective for each 15-day interval of the epidemic by computing R_effective for people who were infected during that 15-day interval.


We are not suggesting that these strategies are practical, nor should school closures alone be used to combat community-wide outbreaks, but it does illustrate how the timing of school closures can reduce or delay the epidemic peak by changing the transmissibility of the pathogen in the population.



\clearpage
# Model description

Corvid is a new agent-based model to simulate SARS-CoV-2 transmission in synthetic communities that represent typical American populations.
It is based on an influenza model ("FluTE"), which is described in detail in doi:10.1371/journal.pcbi.1000656 and is available at https://github.com/dlchao/FluTE.
FluTE was based on the EpiCast model described in Germann et al 2006.
The new model is called "Corvid" and the C++ model code and the analysis scripts used to generate this report are available at https://github.com/dlchao/corvid. We use version 0.3 of Corvid for the runs described in this document.

In brief, Corvid creates communities of about 2000 people, and to simulate larger populations, many such communities are created and connected through commuting patterns.
Synthetic populations for US locations are created by generating enough communities to represent each census tract, with a population size matching the 2000 US Census data, and linking those communities using commuter data from that Census.
Unfortunately, more recent censuses did not offer the commuting data.
Individuals are in 5 age bins: pre-school (0-4y), school-aged (5-18y), young adult (19-29y), older adult (30-64y), and elderly (>65y).
Individuals are generated by populating the communities with families with the household sizes and age distribution drawn from the 1% public use microdata sample of the 2000 Census.
The model runs in discrete half-day time steps, representing day and night.
During the day, people go to institutions (mixing groups) appropriate to their age, and at night they return to their families (where ages mix).
Working-age adults may commute to different census tracts during the day.
Susceptible people become infected when then are in the same setting as infected people. Susceptibles also have a small chance of infection if they are in the same community at the same time as an infectious person.
Upon infection, a person may become infectious starting the next day (regardless of symptom status) and their level of infectiousness can change daily (e.g., exponential, log-normal, etc). A fraction of infectious people become symptomatic after an incubation period, which is specified as a CDF for flexible parameterization (e.g., can be fixed incubation time, normal, Weibull, etc). Symptomatic people may choose to stop going to work or school because of illness. Public health interventions, such as home quarantine or school closures, may be based on the presence of detected symptomatic cases.
Social distancing is implemented by closing settings, like schools, to stop transmission but this is partially offset by individuals going elsewhere.
Much more detail about model structure is in the 2010 publication:
doi:10.1371/journal.pcbi.1000656.
Updates to the model since that publication are described at the end of this document in "Implementation notes".


## Calibration

Corvid generally uses influenza parameters from FluTE, but we had to change the incubation period and infectiousness over time to mimic SARS-CoV-2 transmission.

### Symptomatic fraction, incubation period, shedding (infectiousness) kinetics

We assume that 50% of people infected will become symptomatic, regardless of age.

For the incubation period distribution (duration from infection to showing symptoms), we used the log-normal fit from Lauer et al 2020 who used data from Shenzhen.
They found parameters of log(mean)=1.621 log(sd)=0.418.

```{r incubation, fig.width=4.25, fig.height=3.25, fig.cap="Incubation period distribution and daily infectiousness in Corvid. Black curve is the log-normal distribution of incubation periods from Lauer et al, truncated at day 14. Incubation periods are drawn from this distribution in Corvid. The blue curve is degree of infectiousness (shedding) over time in Corvid, which is a log-normal distribution with a higher mean than but same sd as the incubation distribution. The distribution truncated at day 21. In Corvid, people are infectious (but not highly) on the first day of infection regardless of symptomatic status.", echo=FALSE}
par(mar=c(3.5,3.5,1,1), #bottom, left, top, and right.
    mgp=c(2.0, 0.6, 0))
time <- seq(0,14,0.1)
plot(x=time, y=dlnorm(time, meanlog=1.621, sdlog=0.418), xlim=c(0,21), type="l", lwd=1.5, xlab="days after infection", ylab="density", ylim=c(0,0.25), cex.lab=1.1, cex.axis=1) # from Lauer

time <- seq(0,21,0.1)
lines(x=time, y=dlnorm(time, meanlog=log(exp(1.621)+4), sdlog=0.418), col="blue")
legend("topright", legend=c("incubation period distribution","shedding trajectory"), col=c("black","blue"), lty=rep("solid",2), cex=0.8, inset=0.01)
```

There are reports that viral detection peaks a few days after symptoms.
Wolfel et al found pharyngeal shedding to peak 4 days after symptoms using throat swabs.
Therefore, we just used the incubation period distribution and added 4 to the mean (set the log(mean) of the distribution to log(exp(1.621)+4)).
That makes shedding peak 4 days after the median incubation period.
It might make more sense for shedding to peak right after symptoms appear in an individual instead of independent of symptom onset.
We assume that shedding lasts for 21 days, though the level of shedding is low by the end of the infectious period.
We make asymptomatic individuals half as infectious as symptomatic.
So if a person never becomes symptomatic, infectiousness is simply half that of a symptomatic individual over time. If a person becomes symptomatic on day 5, then for first five days that person is less infectious and after that approximately twice as infectious.
Currently, there is no evidence that asymptomatic people are less infectious than symptomatic, but we assume that the symptoms like sneezing mechanically spread disease.

### Calibrating R_0 in the model

We calibrate R0 by counting the number of people an index case can infect in the model.
This captures not only shedding kinetics, but behavior in response to illness (tendency to stay home) and age-dependent mixing patterns.
We set transmissibility of coronavirus in the model by multiplying contact probabilities by the scalar beta, which we can set in the model.
We derive the relationship between R0 and beta by running the model with different values of beta and counting the number of people an index case infects in a fully susceptible population.
For each value of beta that we test, we infect one randomly selected individual in the population.
We fit a line through the average number of secondary cases for each value of beta tested.
The number of people infected depends on the age of the index case.
For example, if an index case is a school-aged child, the number of secondary cases is higher than for other age groups.
Also, the secondary cases are not representative of the general population -- school-aged children are also over-represented among infectees.
R0 should be defined as the number of secondary cases generated by a "typical" case.
Therefore, we tally the number of secondary cases generated by all index cases across runs to get a better idea of who is more likely to be a "typical" infector.
We weight the relationship based on the proportion of secondary cases in each age bin; we put more weight on the high transmission from school children than the lower transmission from adults in our R0 calculation.

The synthetic population is based on 2000 Census data of the metropolitan Seattle area.
The synthetic population has `r as.integer(sum(tracts.seattle$pop))` residing in `r nrow(tracts.seattle)` census tracts.
Tracts have `r min(tracts.seattle$pop)` to `r as.integer(max(tracts.seattle$pop))` people living in them, and `r min(tracts.seattle$workers)` to `r as.integer(max(tracts.seattle$workers))` people working in them.

```{r map-seattle, fig.width=4, fig.height=4, fig.cap="Map of the census tracts for the synthetic Seattle population. Dot size proportional to resident population. Red circle size is proportional to the number of people who work in each tract.", echo=FALSE}
par(mar=c(3.5,3.5,1,1), #bottom, left, top, and right.
    mgp=c(2.0, 0.6, 0))
plot(x=tractlocs$longitude, y=tractlocs$latitude, col="blue", cex=0.01*sqrt(tracts.seattle$pop[match(tractlocs$fipstract, tracts.seattle$FIPStract)]), pch=19, xlab="longitude", ylab="latitude", asp=1, cex.axis=0.5, cex.lab=0.8)
points(x=tractlocs$longitude, y=tractlocs$latitude, col="red", cex=0.01*sqrt(tracts.seattle$workers[match(tractlocs$fipstract, tracts.seattle$FIPStract)]), pch=1)
```

# Acknowledgments

Thanks to Clinton Collins for helping me run some of these simulations on a decent computer.

\clearpage
# References

* Chao DL, Halloran ME, Obenchain VJ, Longini IM Jr.
  FluTE, a publicly available stochastic influenza epidemic simulation model.
  PLoS Comput Biol. 2010 Jan 29;6(1):e1000656. doi: 10.1371/journal.pcbi.1000656
* Germann TC, Kadau K, Longini IM Jr, Macken CA.
  Mitigation strategies for pandemic influenza in the United States.
  Proc Natl Acad Sci U S A. 2006 Apr 11;103(15):5935-40.
* Germann TC, Gao H, Gambhir M, Plummer A, Biggerstaff M, Reed C, Uzicanin A. School dismissal as a pandemic influenza response: When, where and for how long? Epidemics. 2019 Sep;28:100348. doi: 10.1016/j.epidem.2019.100348.
* Lauer SA, Grantz KH, Bi Q, Jones FK, Zheng Q, Meredith HR, Azman AS, Reich NG, Lessler J.
  The Incubation Period of Coronavirus Disease 2019 (COVID-19) From Publicly Reported Confirmed Cases: Estimation and Application.
  Ann Intern Med. 2020 Mar 10. doi: 10.7326/M20-0504.
* Wolfel R, et al. Virological assessment of hospitalized cases of coronavirus disease 2019.
  MedRxiv 2020. doi: 10.1101/2020.03.05.20030502.
